DeepSphere: a graph-based spherical CNN

### 2019-11-14 main.tex
fix footnote number
typos from JB

% 0.2 Explain why the Laplacian is so important...
By definition, the (spherical) graph convolution is rotation equivariant, if and only if, it commutes with the rotation operator defined as $R(g), g\in SO(3)$: $R(g) f(x) = f\left(g^{-1} x \right)$.
In the context of this work, graph convolution is performed by recursive applications of the graph Laplacian (\eqnref{graph_conv}).
Hence, if $\b{L}$ commutes with the rotation operator, then, by recursion, it will also commute with the convolution $h(\b{L})$.
As a result, to simplify our analysis, we concentrate on the graph Laplacian and $h(\b{L})$ is rotational invariant if and only if
% 0.3 Formulate the problem
\begin{equation} \label{eq:equivariance}
	\b{R}_\V(g) \b{L} \b{f} = \b{L} \b{R}_\V(g) \b{f} \hspace{1cm} \forall f\in F_\V \text{ and } \forall g\in SO(3),
\end{equation}
where $\b{R}_\V(g) = T_\V R(g) T_\V^{-1}$. For practical a evaluation of the equivariance, we define the \textit{normalized equivariance error} for a signal $\b{f}$, and a  rotation $g$ as:
\begin{equation} \label{eq:equivariance error}
	E_{\b{L}}(\b{f}, g) = \left(\frac{ \norm {\b{R}_\V(g) \b{L} \b{f} - \b{L} \b{R}_\V(g) \b{f}} }{\norm {\b{L} \b{f}}}\right)^2,
\end{equation}
More generally for a class of signal $f \in C \subset F_\V$, the \textit{mean equivariance error} defined as
\begin{equation} \label{eq:mean equivariance error}
	\overline E_{\b{L}, C} = \mathbb E_{\b{f}\in C, g\in SO(3)} \ E_{\b{L}}(\b{f}, g),
\end{equation}
represents well the overall equivariance error.
In practice, we analyse this error with respect of the frequency by setting the set $C$ to functions $f$ made of single degree spherical harmonic.
The expected value is obtained by averaging over a finite number of random functions and random rotations.

@@

% Nevertheless, it is important to keep in mind the limits of Theorem \ref{theo:pointwise convergence for a regular sampling}.
% First, it is \emph{not} a proof of equivariance of the graph convolution for a specific sampling. To construct such a proof, a stronger convergence result is needed, for example spectral convergence (instead of pointwise).
% \todo{Can we intuitively write what pointwise convergence means, and how it is more limited than spectral or uniform convergence?}
We note that in the random sampling case, the graph laplacian does \emph{not} converge uniformly in general \cite[Sec. 3]{belkin2007convergence}, but does for a class of Lipschitz functions \cite[Sec. 6]{belkin2005towards}.
Second, while do not have a formal proof for it, we strongly believe that the HEALPix sampling does satisfy the hypothesis $d^{(n)}\leq \frac{C}{n^\alpha}, \ \alpha\in (0,\linefrac{1}{2}]$ with $\alpha$ very close or equal to $\linefrac{1}{2}$. The empirical results discussed in the next paragraph also points in this direction. This is further discussed in Appendix \ref{sec: appendix: proof of theorem}.
% 3) Show experimental results

@@

% 3.1) explain what the figures tell us
For 8 neighbors, selecting the optimal value of $t$ already leads to an improvement compared to \citet{perraudin2019deepspherecosmo}, whose graph also was constructed to have at maximum 8 neighbors.
With higher the number of neighbors, the equivariance error decreases by more than one order of magnitude.
Furthermore, increasing the graph resolution ($N_{side}$) leads to an improvement in the high frequencies as well.
Increasing the graph resolution of the graph of \citet{khasanova2017sphericalcnn} does not seem to improve its equivariance. We note that in our numerical test, this graph was perfectly equivariant to rotations along the horizontal and vertical axis as it should be by design. Nevertheless, it is simply not equivariant along the last rotation. Such equivariance properties might be sufficient for analysing omnidirectional images (that are naturally aligned on the north-south axis).

@@

\begin{table}
    \centering
	%\begin{tabular}{l cc r rr}
	\begin{tabular}{l cc r rS[table-format=2.0]}
		\toprule
		& \multicolumn{2}{c}{performance} & \multicolumn{1}{c}{size} & \multicolumn{2}{c}{speed} \\
		\cmidrule(lr){2-3} \cmidrule(lr){4-4} \cmidrule(lr){5-6}
		& F1 & mAP & params & \multicolumn{1}{c}{inference} & \multicolumn{1}{c}{training} \\
		\midrule
		% $SO(3)$ vs $S^2$ vs graph
		\citet{cohen2018sphericalcnn} ($b=128$) & - & 67.6 & 1400\,k & 38.0\,ms & 50\,h \\
		\citet{cohen2018sphericalcnn} (simplified,\protect\footnotemark $b=64$) & 78.9 & 66.5 & 400\,k & 12.0\,ms & 32\,h \\
		\citet{esteves2018sphericalcnn} ($b=64$) & 79.4 & 68.5 & 500\,k & 9.8\,ms & 3\,h \\
		DeepSphere (equiangular $b=64$) & 79.4 & 66.5 & 190\,k & 0.9\,ms & 50\,m \\
		DeepSphere (HEALPix $N_{side}=32$) & 80.7 & 68.6 & 190\,k & 0.9\,ms & 50\,m \\
		\bottomrule
	\end{tabular}
    \caption{
		Results on SHREC'17 (3D shapes). DeepSphere achieves similar performance at a much lower cost, suggesting that anisotropic filters are an unnecessary price to pay.
		% F1-score computed with sklearn, mAP from the official script of the competition.
		% inference speed = time for a single instance to do a single training pass
		% training speed = time for the neural net to train to peak performance
	}
    \label{tab:shrec17}
\end{table}
\footnotetext[5]{As implemented in \url{https://github.com/jonas-koehler/s2cnn}.}

>>>>

% 0.2 Explain why the Laplacian is so important...
By definition, the (spherical) graph convolution is rotation equivariant, if and only if, it commutes with the rotation operator defined as $R(g), g\in SO(3)$: $R(g) f(x) = f\left(g^{-1} x \right)$.
In the context of this work, graph convolution is performed by recursive applications of the graph Laplacian (\eqnref{graph_conv}).
Hence, if $\b{L}$ commutes with the rotation operator, then, by recursion, it will also commute with the convolution $h(\b{L})$.
As a result, to simplify our analysis, we concentrate on the graph Laplacian and $h(\b{L})$ is rotational invariant if and only if
% 0.3 Formulate the problem
\begin{equation} \label{eq:equivariance}
	\b{R}_\V(g) \b{L} \b{f} = \b{L} \b{R}_\V(g) \b{f} \hspace{1cm} \forall f\in F_\V \text{ and } \forall g\in SO(3),
\end{equation}
where $\b{R}_\V(g) = T_\V R(g) T_\V^{-1}$. For an empirical evaluation of the equivariance, we define the \textit{normalized equivariance error} for a signal $\b{f}$, and a  rotation $g$ as:
\begin{equation} \label{eq:equivariance error}
	E_{\b{L}}(\b{f}, g) = \left(\frac{ \norm {\b{R}_\V(g) \b{L} \b{f} - \b{L} \b{R}_\V(g) \b{f}} }{\norm {\b{L} \b{f}}}\right)^2,
\end{equation}
More generally for a class of signal $f \in C \subset F_\V$, the \textit{mean equivariance error} defined as
\begin{equation} \label{eq:mean equivariance error}
	\overline E_{\b{L}, C} = \mathbb E_{\b{f}\in C, g\in SO(3)} \ E_{\b{L}}(\b{f}, g),
\end{equation}
represents well the overall equivariance error.
In practice, we analyse this error with respect of the frequency by setting the set $C$ to functions $f$ made of single degree spherical harmonic.
The expected value is obtained by averaging over a finite number of random functions and random rotations.

@@

% Nevertheless, it is important to keep in mind the limits of Theorem \ref{theo:pointwise convergence for a regular sampling}.
% First, it is \emph{not} a proof of equivariance of the graph convolution for a specific sampling. To construct such a proof, a stronger convergence result is needed, for example spectral convergence (instead of pointwise).
% \todo{Can we intuitively write what pointwise convergence means, and how it is more limited than spectral or uniform convergence?}
We note that in the random sampling case, the graph laplacian does \emph{not} converge uniformly in general \cite[Sec. 3]{belkin2007convergence}, but does for a class of Lipschitz functions \cite[Sec. 6]{belkin2005towards}.
Second, while we do not have a formal proof for it, we strongly believe that the HEALPix sampling does satisfy the hypothesis $d^{(n)}\leq \frac{C}{n^\alpha}, \ \alpha\in (0,\linefrac{1}{2}]$ with $\alpha$ very close or equal to $\linefrac{1}{2}$. The empirical results discussed in the next paragraph also points in this direction. This is further discussed in Appendix \ref{sec: appendix: proof of theorem}.
% 3) Show experimental results

@@

% 3.1) explain what the figures tell us
For 8 neighbors, selecting the optimal value of $t$ already leads to an improvement compared to \citet{perraudin2019deepspherecosmo}, whose graph also was constructed to have at most 8 neighbors.
With higher the number of neighbors, the equivariance error decreases by more than one order of magnitude.
Furthermore, increasing the graph resolution ($N_{side}$) leads to an improvement in the high frequencies as well.
Increasing the graph resolution of the graph of \citet{khasanova2017sphericalcnn} does not seem to improve its equivariance. We note that in our numerical test, this graph was perfectly equivariant to rotations along the horizontal and vertical axis as it should be by design. Nevertheless, it is simply not equivariant along the last rotation. Such equivariance properties might be sufficient for analysing omnidirectional images (that are naturally aligned on the north-south axis).

@@

\begin{table}
    \centering
	%\begin{tabular}{l cc r rr}
	\begin{tabular}{l cc r rS[table-format=2.0]}
		\toprule
		& \multicolumn{2}{c}{performance} & \multicolumn{1}{c}{size} & \multicolumn{2}{c}{speed} \\
		\cmidrule(lr){2-3} \cmidrule(lr){4-4} \cmidrule(lr){5-6}
		& F1 & mAP & params & \multicolumn{1}{c}{inference} & \multicolumn{1}{c}{training} \\
		\midrule
		% $SO(3)$ vs $S^2$ vs graph
		\citet{cohen2018sphericalcnn} ($b=128$) & - & 67.6 & 1400\,k & 38.0\,ms & 50\,h \\
		\citet{cohen2018sphericalcnn} (simplified,\protect\footnotemark $b=64$) & 78.9 & 66.5 & 400\,k & 12.0\,ms & 32\,h \\
		\citet{esteves2018sphericalcnn} ($b=64$) & 79.4 & 68.5 & 500\,k & 9.8\,ms & 3\,h \\
		DeepSphere (equiangular $b=64$) & 79.4 & 66.5 & 190\,k & 0.9\,ms & 50\,m \\
		DeepSphere (HEALPix $N_{side}=32$) & 80.7 & 68.6 & 190\,k & 0.9\,ms & 50\,m \\
		\bottomrule
	\end{tabular}
    \caption{
		Results on SHREC'17 (3D shapes). DeepSphere achieves similar performance at a much lower cost, suggesting that anisotropic filters are an unnecessary price to pay.
		% F1-score computed with sklearn, mAP from the official script of the competition.
		% inference speed = time for a single instance to do a single training pass
		% training speed = time for the neural net to train to peak performance
	}
    \label{tab:shrec17}
\end{table}
\footnotetext[7]{As implemented in \url{https://github.com/jonas-koehler/s2cnn}.}
